Estimating Substitution Using Text Embeddings: \\ Evidence from the Movie Industry

### 2024-09-18 draft/draft.tex
Minor draft polish

Given the above,our demand model assumes characteristics are exogeneous. If instead they were chosen as equilibrium objects, then the coefficients on proximity are no longer causal: the impact of locating near a close competitor becomes confounded with latent demand characteristics driving the \emph{choice} to locate near a close competitor. However, the existance of twin movies suggests this is not a major concern.

@@

Aggregating over all consumers, we can express the demand for movie $i$ in week $t$ as a function $\phi$ of the movie's appeal $\delta_{it}$ and the appeal of its competitors $\delta_{-it}$. Again due to the uniform pricing scheme, we can use revenues as a proxy for quantity, since quantity is simply a constant scalar transformation of revenue. Finally, demand for movies varies idiosyncratically by weekend; for example, there is much more demand for movies on Christmas Day than on a random weekend in sunny July. Thus, I include a fixed effect $\alpha_t$ to capture week-by-week variation. Thus, I can write the demand for movie $i$ in week $t$ as:

@@

As a check for the quality of the embeddings, I perform t-Stochastic Neighbor Embedding (t-SNE) on the embeddings to reduce them to two dimensions and highlight clustering. For each movie, I identify the first genre listed on the TMDB page.Embeddings do cluster by genre, as visible in figure \ref{fig:tsne}. Generally, we see action and adventure movies grouped in the top-right quadrant, science fiction and fantasy in the bottom-right quadrant, and romance and comedies on the left; that is, it seems as through the embeddings are not just clustering \emph{within} a genre, but also grouping \emph{similar genres}.

@@

Nearest neighbors to some common movies are visible in table \ref{tab:cosine}. Just by inspection, it is clear this similarity score is capturing \emph{something}: the closest movies to \emph{The Notebook} are other Nicholas Sparks book adaptations, and the closest movies to \emph{A Bug's Life} are \emph{The Ant Bully} and \emph{Antz}. Moreover, the closest neighbors to \emph{Frozen} and \emph{Avengers: Endgame} are in the same franchise: \emph{Frozen II} and \emph{Avengers: Infinity War} respectively. Thus, I feel comfortable using these embeddings to capture similarity between movies.

@@

\textcite{harper2016ATIIS} provide a public-use dataset of real movie reviews (``MovieLens''); it contains more than 25 million 1-to-5 star reviews of movies by users. Moreover, movies are tagged with their TMDB ID; that is, the same ID I use to collect movie descriptions. Thus, for any given pair of movies, I can compute similarity under both the text-based and collaborative filtering methods. If the two measures are correlated, I can be more confident that the patterns I see in the text-based model are not coincidental. I validate my text-based model by comparing it against two frontier collaborative filtering algorithms: topic-specific PageRank and UV decomposition.

@@

>>>>

Given the above, our demand model assumes characteristics are exogeneous. If instead they were chosen as equilibrium objects, then the coefficients on proximity are no longer causal: the impact of locating near a close competitor becomes confounded with latent demand characteristics driving the \emph{choice} to locate near a close competitor. However, the existance of twin movies suggests this is not a major concern.

@@

Aggregating over all consumers, we can express the demand for movie $i$ in week $t$ as a function $\phi$ of the movie's appeal $\delta_{it}$ and the appeal of its competitors $\delta_{-it}$. Again due to the uniform pricing scheme, we can use revenues as a proxy for quantity, since quantity is simply a constant scalar transformation of revenue. Finally, demand for movies varies idiosyncratically by weekend; for example, there is much more demand for movies on Christmas Day than on a random weekend in sunny July. Thus, I include a fixed effect $\alpha_t$ to capture week-by-week variation. I can therefore write the demand for movie $i$ in week $t$ as:

@@

As a check for the quality of the embeddings, I perform t-Stochastic Neighbor Embedding (t-SNE) on the embeddings to reduce them to two dimensions and highlight clustering. For each movie, I identify the first genre listed on the TMDB page. Embeddings do cluster by genre, as visible in figure \ref{fig:tsne}. Generally, we see action and adventure movies grouped in the top-right quadrant, science fiction and fantasy in the bottom-right quadrant, and romance and comedies on the left; that is, it seems as through the embeddings are not just clustering \emph{within} a genre, but also grouping \emph{similar genres}.

@@

Nearest neighbors to some popular movies are visible in table \ref{tab:cosine}. Just by inspection, it is clear this similarity score is capturing \emph{something}: the closest movies to \emph{The Notebook} are other Nicholas Sparks book adaptations, and the closest movies to \emph{A Bug's Life} are \emph{The Ant Bully} and \emph{Antz}. Moreover, the closest neighbors to \emph{Frozen} and \emph{Avengers: Endgame} are in the same franchise: \emph{Frozen II} and \emph{Avengers: Infinity War} respectively. Thus, I feel comfortable using these embeddings to capture similarity between movies.

@@

\textcite{harper2016ATIIS} provide a public-use dataset of real movie reviews (``MovieLens''); it contains more than 25 million 1-to-5 star reviews of movies by users. Moreover, movies are tagged with their TMDB ID; that is, the same ID I use to collect movie descriptions. Thus, for any given pair of movies, I can compute similarity under both the text-based and collaborative filtering methods. If the two measures are correlated, I can be more confident that the patterns I see in the text-based model are not coincidental. I validate my text-based model by comparing it against two frontier collaborative filtering algorithms: UV decomposition and topic-specific PageRank.

@@

% \section*{Acknowledgments}
% Thank you to Liran Einav and Shoshana Vasserman for their support both in and after IO III. Thanks also to Arun Chandrasekhar, Nick Bloom, Janet Stefanov, Nick Scott-Hearn, and Lauren Harris for their feedback on this project, as well as the rest of the 2024 IO III class.
